---
title: "2015 CT High School Analysis"
author: "Chuck Stietzel"
date: "11/24/2017"
output: ioslides_presentation
---

<style type="text/css">
pre {
  font-size: 12px;
  line-height: 16px;
  padding: 0px 0px 10px 60px;
}
</style>

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(knitr)
opts_chunk$set(echo = FALSE)

```
## High School Analysis
- There is a great deal of interest in improving the performance of public education. This application employs a simplistic linear model using data published by the State of Connecticut to illustrate potential relationships between outcomes and input variables.

-  App Features:
    - Select from different performance outcomes
    - Compare outcome to one of three variables using a scatterplot and histograms
    - Examine a linear model fitted to the data 
    - View all of the study data in a user-friendly way

## CT High School Data
- Data obtained from State of Connecticut's [EdSight](http://edsight.ct.gov/SASPortal/main.do) web portal
- As an exploratory analysis Absenteeism and Pupil Spending were plotted
- Enrollment was bucketed
    - Small : less than 500
    - Medium: between 500 and 100
    - Large: more than 1000
- Per pupil spending was obtained at the district level
- Absenteeism obtained at the school level

```{r dataprep, warning=FALSE, message=FALSE, echo=FALSE}
## Get School Data
schools <- read_csv("HSApp/data/school_w_loc.csv")

## Get School Enrollment Data
enrollment <- read_csv("HSApp/data/enrollmentSingleYear14-15.csv", skip = 6, na = "*",
                       col_names  = c("DistrictName", "DistrictCode", "SchoolName", "SchoolCode", "NativeAmerican", 
                                      "Asian", "Black", "Hispanic", "Islander", "Mixed", "White", "Total")) %>% 
  
  ## Remove all '=' and '"' symbols from file
  mutate_at(vars(DistrictCode, SchoolCode), function(x) as.integer(gsub("[=\"]+", "", x)))

## Get Absenteeism data
absent <- read_csv("HSApp/data/chronicAbsenteeism14-15.csv", skip = 4, na = "*",
                   col_names  = c("DistrictName", "DistrictCode", "SchoolName", "SchoolCode", "PctAbsent")) %>% 
  
  ## Remove all '=' and '"' symbols from file
  mutate_at(vars(DistrictCode, SchoolCode), function(x) as.integer(gsub("[=\"]+", "", x))) %>%
  select(SchoolCode, PctAbsent)

## Get Per Pupil Spending data
pupilspend <- read_csv("HSApp/data/perPupilExpeditures14-15.csv", skip = 5, na = "N/A", 
                       col_names  = c("DistrictName", "TeachingStaff", "TeachingSupplies", "MediaServices", "StudentSupport", 
                                      "Administration", "Facilities", "Transportation", "Other", "TotalPP")) %>% 

    ## Remove all '=' and '"' symbols from file
  mutate_at(vars(DistrictName), function(x) gsub("[=\"]+", "", x))

hsdata <- semi_join(enrollment, schools, by = c("SchoolCode" = "OrganizationCode")) %>%
  inner_join(absent, by = "SchoolCode") %>%
  inner_join(pupilspend, by = "DistrictName") %>%

    ## Add School Size Factor
  mutate(Size = cut(Total, breaks=c(0,500,1000,3000), labels = c("Small", "Medium", "Large")))
```

## District Spending vs Absentee Rate
```{r plot, warning=FALSE, message=FALSE, echo=TRUE}
y <- list(title = "Per Pupil Spending ($)"); x <- list(title = "School Days Missed (%)")
plot_ly(hsdata, x = ~PctAbsent, y = ~TotalPP, type = "scatter", color = ~Size, size = ~Total, text = ~SchoolName) %>%
  layout(autosize = F, width = 750, height = 450, xaxis = x, yaxis = y, title = "Academic Year 2014/15")
```

## 2015 High School Performance App {.smaller}
- Inputs are averages of academic years 2011/12 to 2014/15
    - <font size="4">Absenteeism and Enrollment obtained at the school level</font>
    - <font size="4">Per Pupil Spending obtained at the district level</font>

<BR>

- Student performance outcomes are for the Class of 2015 cohort
    - <font size="4">*Graduation Rate*: % of students that graduated in four years</font>
    - <font size="4">*College Entry Rate*: % of students that entered a post secondary school</font>
  
<BR>
  
- Links
    - <font size="4">HSApp application on shinyapps.io: [HSApp](https://cstietzel.shinyapps.io/HSApp/)</font>
    - <font size="4">Link to pitch book on Github: [HSAppSlides.html](https://cstietzel.github.io/ds9-dataproducts/wk4-app/HSAppSlides.html)</font>
    - <font size="4">Code and Rmd on Github: [HSApp Code](https://github.com/cstietzel/ds9-dataproducts/tree/gh-pages/wk4-app)</font>
    
    
